---
title: "p8105_hw6_lcc2181"
output: github_document
---

This code chunk loads packages to complete Homework 6, adds formatting for plots, and sets a seed to ensure reproducibility.

```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(purrr)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

set.seed(1)
```

## Problem 1

Loading and cleaning data for regression analysis

* Importing data
* Converting categorical variables into factors
* Checking for missing data

```{r, message=FALSE}
birthweight_df =
  read_csv(file = "./data/birthweight.csv") %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  )

birthweight_df %>% 
  is.na.data.frame() %>% 
  sum()
```

Proposing a regression model for birthweight:

* The model `fit` is based on a hypothesized structured for factors associated with birthweight
* Specifically, the predictors included in the model were based on the following factors which may underly birthweight:
baby's sex, baby's length at birth, baby's head circumference, mother's age, mother's race, gestational age, and number of cigarettes smoked per day during pregnancy

```{r}
fit = lm(bwt ~ babysex + blength + bhead + momage + mrace + gaweeks + smoken, data = birthweight_df)
```

Plotting model residuals against fitted values:

```{r}
birthweight_df %>% 
  add_predictions(fit) %>% 
  add_residuals(fit) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  labs(x = "Fitted Values", y = "Residuals")
```

Comparing model `fit` to two other models in terms of cross-validated prediction error:

* `fit_1`: Model using length at birth and gestational age as predictors
* `fit_2`: Model using head circumference, length at birth and sex (as well as all interactions between these) as predictors

```{r}
cv_df =
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>% 
  mutate(
    fit = map(train, ~lm(bwt ~ babysex + blength + bhead + momage + mrace + gaweeks + smoken, 
          data = .x)),
    fit_1 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    fit_2 = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex 
          + blength * babysex + bhead * blength * babysex, data = .x)),
  ) %>% 
  mutate(
    rmse_fit = map2_dbl(fit, test, ~rmse(model = .x, data = .y)),
    rmse_fit_1 = map2_dbl(fit_1, test, ~rmse(model = .x, data = .y)),
    rmse_fit_2 = map2_dbl(fit_2, test, ~rmse(model = .x, data = .y))
  )
```

Plotting distribution of RMSE for each model:

```{r}
cv_df %>% 
  select(rmse_fit, rmse_fit_1, rmse_fit_2) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

Based on this plot, the model I created (`fit`) performs better than the other two models (`fit_1` and `fit_2`).

## Problem 2

The code chunk below downloads 2017 Central Park weather data.

```{r, message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Drawing 5,000 bootstrap samples, and for each bootstrap sample, producing estimates of R-squared and log(beta0 * beta1)

```{r, message=FALSE}
bootstrap_results = 
  weather_df %>% 
  bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results_1 = map(models, broom::glance),
    results_2 = map(models, broom::tidy)
  ) %>% 
  select(results_1, results_2) %>% 
  unnest(results_1, results_2) %>% 
  select(r.squared, term, estimate) %>% 
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(log_beta0_beta1 = log(intercept*tmin)) %>% 
  select(-intercept, -tmin)
```

Plotting estimates for R-squared and log(beta0 * beta1)

```{r}
plot_1 =
  bootstrap_results %>% 
  ggplot(aes(x = r_squared)) +
  geom_density()

plot_2 =
  bootstrap_results %>% 
  ggplot(aes(x = log_beta0_beta1)) +
  geom_density() +
  labs(x = "log(beta0 * beta1)")

plot_1 + plot_2
```

The distributions of R-squared and log(beta0 *beta1) both appear to be relatively normally distributed, although both are slightly left-tailed.

The following code chunk identifies 2.5% and 97.5% quantiles for R-squared and log(beta0 * beta1):

```{r}
bootstrap_results %>%
  summarize(
    ci_rsquared_lower = quantile(r_squared, 0.025),
    ci_rsquared_upper = quantile(r_squared, 0.975)
  )

bootstrap_results %>%
  summarize(
    ci_log_beta0_beta1_lower = quantile(log_beta0_beta1, 0.025),
    ci_log_beta0_beta1_upper = quantile(log_beta0_beta1, 0.975)
  )
```

Based on this output, the 95% CIs for R-squared and log(beta0 * beta1) are:

* 95% CI for R-squared: (0.89, 0.93)
* 95% CI for log(beta0 * beta1): (1.96, 2.06)